\subsection{Motivations for developing OHD-GAN}
The authors cite a wide range of potential applications for generative models of OHD. While some of these goals are over-optimistic and have yet to be realized, they paint an encouraging picture for the value of synthetic OHD and the transformative effect it could have on healthcare initiatives and scientific progress. We briefly describe the four prevailing themes in Sec.\ref{sec:augmentation} data augmentation, Sec.\ref{sec:access_privacy} privacy and accessibility, Sec.\ref{sec:precision_med} precision medicine and Sec.\ref{sec:models_twins} modelling. 

    \subsubsection{Data augmentation}\label{sec:augmentation}
    Data augmentation is mentioned in nearly all publications. Most commonly, synthetic data can improve generalization in predictive algorithms by providing additional information about the real data distribution \cite{Wang_2019,Che_2017,Yoon2018-dm, Yoon2018-mo}. Similarly, the application of GANs to domain translation and semi-supervised training approaches could support predictive tasks in healthcare that lack data with accurate labels, paired samples, or present class imbalance \cite{Che_2017,mcdermott2018semi}. 

    \subsubsection{Enhancing privacy and increasing accessibility}\label{sec:access_privacy}
    SD is seen as the key to unlocking the value of OHD, which is currently locked in due to privacy concerns. Preserving privacy can broadly be described as reducing the risk of re-identification attacks to an acceptable level. Many studies noted that highly restricted access to OHD is hindering machine learning, and more generally scientific progress \cite{Beaulieu-Jones2019-ct, Baowaly_2019,Che_2017,esteban2017real,Fisher2019}. Due to its artificial nature, SD is proposed as a means to forgo data use agreements, while potentially providing greater privacy guarantees and reducing the loss of utility \cite{Beaulieu-Jones2019-ct, Baowaly_2019,esteban2017real,Fisher2019,walsh2020generating}. Overall, enabling access to greater variety, quality and quantity of OHD could have positive effects in a wide range of fields, such as software development, education, and training of medical professionals. 
    
    \subsubsection{Enabling precision medicine}\label{sec:precision_med}
    The ability to conduct simulations of disease progression for individual patients could have transformative impacts on healthcare. Generative models conditioned on a patient's baseline state could help inform clinical decision making by quantifying disease progression \cite{walsh2020generating, Fisher2019}. Additionally, stochastic simulations of individual patient profiles could help quantify risk at an unprecedented level of granularity \cite{Fisher2019}. Predicting patient-specific responses to drugs is still a new field of research, a problem known as Individualized Treatment Effects (ITE). The task of estimating ITEs is persistently hampered by the lack or paired samples, or counterfactuals \cite{Yoon2018-mo, chu2019treatment}. In regards to these issues, GANs have shown the ability for domain translation, mapping a sample from its to original class to the paired equivalent. This includes bidirectional transformations, in addition to the possibility of learning from very few paired samples, or even achieving better performance in the absence of paired samples \cite{Wolterink2017DeepMT}.
    
    \subsubsection{From patient and disease models to digital twins}\label{sec:models_twins}
    Realistic synthetic data implies a model that approximates the process that generated the real information \cite{esteban2017real}. Achieving models of significant complexity would open up new simulation possibilities for developing predictive systems and methods. In clinical research, such models could help quantify cause and effect, simulate different study designs, provide control samples or more generally give us a better understanding of disease progression in relation to initial conditions \cite{Fisher2019, yahi2017generative, walsh2020generating}. Pushing the aspect of simulation further, the concept of "digital twins" represents in a way the ultimate realization of personalized medicine. A common practice in industrial sectors is high-fidelity virtual representations, or long-term simulations, of physical assets that grant a comprehensive understanding of the workings, behavior and life-cycle. Their state is continuously updated from theoretical data, real data, streaming IoT indicators and conditional synthetic data. In a position paper, Angulo et al. draw the parallels of this technique with the current needs in healthcare and the emergence of the necessary technologies for the proposal they bring forward \cite{angulo2019towards,Angulo_2020}. Notably, the rapid adoption of wearables that are continuously monitoring people's physiological state. Through continuous lifelong learning, patient models inform the decisions of medical professionals, but also enable testing research hypotheses. In their proposal, GANs are an essential component of the ecosystem to ensure patient privacy and to provide bootstrap data. Fisher et al. employ the term to describe their method \cite{walsh2020generating}.